Free Probability Analysis of Capacity in Relay Networks

### 2020-04-08 documentation/systemModel/systemModel.tex
Some changes to code but mostly working on writing work

\begin{gather*}
\mQ_{YY} =
\mH\mX\mX^H\mH^H + \mH_I \mX_I\mX_I^H\mH_I^H + \mW\mW^H.
\end{gather*}
\subsection{Free Probability Analysis}\label{decont_fpt}

\begin{equation}
\mathbf{D} = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
\end{equation}
with $\alpha_k \in \mathbb{R}$ is considered. Given the S-transform of the *self-adjoint matrices $\mathbf{B}_k $ and $\mathbf{C}_k$ (\cite{nica2006lectures}) we can first use the product property of the S-transform to find the Stieltjes transform of each term in the sum $\mathbf{B}_k \mathbf{C}_k$. Given the Stieltjes transform of each term, we then find the R-transform of each term $\mathbf{B}_k \mathbf{C}_k$ and exploiting the additive property of the R-transform, we find the R-transform of $\mathbf{D}$. Inverting the R-transform to get the Stieltjes transform of $\mathbf{D}$ which we can evaluate to find the AED.
\par
Note that the resulting Stieltjes transform is given as a fixed point equation which is then evaluated for each value of of interest to find the corresponding value of the AED.

@@

Noting that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes

@@

In this case, the composite channel is given by 
\begin{equation}
\mathbf{H}_{Total} = \mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
\end{equation}
with covariance matrix 
\begin{equation}\label{polynomial_terms}
\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
\end{equation}
Using the rotational property of the trace as in the previous case
We can equivalently look for the AED of the matrix 
\begin{equation}
\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
\end{equation}
Iterating this procedure and noting that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, this is exactly the case from \cite[Section 4.10]{muller2013applications}.

@@

@@

\chapter{Next Steps}
\section{Independent TODO items}
\begin{itemize}
\item 
	Show why the individual channel matrices are free as assumed in decontamination work.
\item
	Work on example of how to find AED given a fixed point equation for the Stieltjes transform.
\end{itemize}
\section{Questions}
\begin{itemize}
\item
	In Pilot Decontamination paper, the final reference refers to the case in which each variable is self-adjoint.
	Are the non-similar terms canceling out?
\item 
	Just having the IRS present should give us a boost, do we need to actually consider selecting the phases?

>>>>

\begin{gather*}
\mQ_{YY} =
\mH\mX\mX^H\mH^H + \mH_I \mX_I\mX_I^H\mH_I^H + \mW\mW^H.
\end{gather*}
\subsection{Free Probability Analysis}\label{decont_fpt}
Consider a channel describe by the matrix sum
\begin{equation}
\mathbf{D} = \sum_{k=1}^{K}\mathbf{D}_k = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
\end{equation}
with $\alpha_k \in \mathbb{R}$. Ultimately the goal is to find the AED of the covariance matrix $C_{DD^H}$ using the Stieltjes transform  $G_{C_{DD^H}}(s)$. In order to do this the following steps will be used.

Given the S-transform of the *self-adjoint matrices $\mathbf{B}_k $ and $\mathbf{C}_k$ (\cite{nica2006lectures}) we can first use the product property of the S-transform to find the Stieltjes transform of each term in the sum $\mathbf{B}_k \mathbf{C}_k$. Given the Stieltjes transform of each term, we then find the R-transform of each term $\mathbf{B}_k \mathbf{C}_k$ and exploiting the additive property of the R-transform, we find the R-transform of $\mathbf{D}$. Inverting the R-transform to get the Stieltjes transform of $\mathbf{D}$ which we can evaluate to find the AED.
\par
Note that the resulting Stieltjes transform is given as a fixed point equation which is then evaluated for each value of of interest to find the corresponding value of the AED.

@@

Assuming cross terms cancel and that $\boldsymbol{\Theta}^{1}\boldsymbol{\Theta}^{1*} = \mathbf{I}$ this becomes

@@

In this case, the composite channel is given by 
\begin{equation}
\mathbf{H}_{Total} = \mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}
\end{equation}
with covariance matrix 
\begin{equation}\label{polynomial_terms}
\mathbf{Q}_{bb} = \mathbf{H}_{Total}\mathbf{H}_{Total}^H = [\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}\mathbf{H}^{1}]^H
\end{equation}
Using the rotational property of the trace as in the previous case
We can equivalently look for the AED of the matrix 
\begin{equation}
\tilde{\mathbf{Q}}_{bb} = [\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}][\mathbf{G}^{B}\boldsymbol{\Theta}^{N}\mathbf{H}^{N} \cdots \boldsymbol{\Theta}^{1}]\mathbf{H}^{1}\mathbf{H}^{1H}].
\end{equation}
Assuming cross terms cancel and that $\boldsymbol{\Theta}^{n}\boldsymbol{\Theta}^{n*} = \mathbf{I}$ $\forall n \in [1\cdots N]$, we iterate the above procedure just as in \cite[Section 4.10]{muller2013applications}.

@@

\section{Solving } 
\subsection{General Approach }

Using the model from the decontamination work we say the total channel is described by 
\begin{equation}
\mathbf{D} = \sum_{k=1}^{K}\mathbf{D}_k = \sum_{k=1}^{K} \alpha_k \mathbf{B}_k \mathbf{C}_k
\end{equation}
We now consider how to ultimately arrive at an expression for the AED of $DD^H$
\begin{itemize}

\item 
	We first find the Stieltjes transform of each term in the total channel $G_{D_kD_k^H}(s)$.
\item
	Now we use a key property of the Stieljes transform
	\begin{equation}
	G_{\lambda^2}(s) = \frac{G_{\lambda}(\sqrt{s})- G_{\lambda}(-\sqrt{s})}{2\sqrt{s}}
	\end{equation}
	to find the Stieltjes transform of the symmetrized singular value decomposition of $D_k$.
\item
	Taking the R-transform, $\tilde{R}_{D_k}(w)$ of the result, we then use the free additive convolution property in order to 
	find the R-transform of the entire equivalent channel $\tilde{R}_{D}(w)$. 
\item
	Using the key property of the Stieltjes transform again we can find the the R-transform of the 
	covariance matrix of the symmetrized channel, $\tilde{R}_{DD^H}(w)$ . We can then find $\tilde{G}_{DD^H}(s)$ and 
	use the inversion to find the symmetrized AED.

\item 
	Noting that a covariance matrix is positive semi-definite, all eigenvalues are are non-negative. Therefore,
	the symmetrized AED is equal to the true AED when evaluated for non-negative values.

\end{itemize}

\subsection{Solving for $G_{D_kD_k^H}(s)$ } 
For the model presented we have 
\begin{equation}
D_k = G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1
\end{equation}
and 
\begin{equation}
C_N = D_kD_k^H = [G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1]
[H_1 \boldsymbol{\Theta}_1 \cdots H_N \boldsymbol{\Theta}_N G_B]
\end{equation}
\begin{itemize}
\item 
	Because commuting matrix multiplication does not change the non-zero eigen value distribution, we consider 
	the matrix 
\begin{equation}
\tilde{C}_N = [\boldsymbol{\Theta}_1^H \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H][G_B \boldsymbol{\Theta}_N  \cdots \boldsymbol{\Theta}_1 H_1 H_1^H]
\end{equation}

\item 
	We use the following key property of the Stieltjes transform and S-Transform. 
\begin{equation}
S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N})
\end{equation}

\item 
	Iterating this procedure, we can see that ultimately the phase matrices will all cancel out.

\end{itemize}

@@

\chapter{Next Steps}
\section{Independent TODO items}
\begin{itemize}
\item Key is to look at 
\end{itemize}
\section{Questions}
\begin{itemize}
\item
	In Pilot Decontamination paper, the final reference refers to the case in which each variable is self-adjoint.
	Are the non-similar terms canceling out?
\item 
	Just having the IRS present should give us a boost, do we need to actually consider selecting the phases?
